Stock enhancement undermines community stability

### 2022-02-01 rmd/manuscript.Rmd
edit: manuscript

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for the fitness difference of captive-bred individuals due to genetic effects and/or plasticity [@weberInteractionsHatcheryWild2003; @snyderLimitationsCaptiveBreeding1996; @arakiHatcheryStockingHelp2010] when considering the reproductive contribution of released individuals to the next generation. Without loss of generality, we assumed constant $K_i$ ($K_i = K$) and $R_t$ ($R_t = R$) across species and time, respectively.

@@

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the demographic benefit of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the long-term average of the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar in a range of ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figure S1-6**).

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon *Oncorhynchus masou masou* (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. Masu salmon is one of Japan's most valuable freshwater resources, and over 10 million hatchery fish are released annually [@kitadaLessonsJapanMarine2020]. In the protected watersheds (**Figure 1C and 1D**), a long-term program exists to monitor stream fish communities along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see **Materials and Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

@@

Competition is a likely mechanism, as our theory assumes. Field experiments confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage, typically one year [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @weberInteractionsHatcheryWild2003]. Importantly, socioeconomic factors control the number of releases (e.g., budget allocation). As such, released fish are probably "excessive" given the density-dependent regulation of recipient communities and may cause resource competition that would otherwise not exist.

@@

We should prioritize habitat preservation and restoration with a broader scope of biodiversity conservation. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of stock enhancement. Protected areas and environmental restoration are promising tools to conserve biodiversity, and the smart spatial design is key to achieving successful conservation [@koningNetworkGrassrootsReserves2020; @akasakaSmartProtectedArea2017]. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@akasakaSmartProtectedArea2017; @teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020]. Socioeconomic factors such as governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options for the sustainable future exist before blindly accepting stock enhancement. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the stock enhancement will never be effective but impairs biodiversity.

# Data availability

@@

>>>>

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for the fitness difference of captive-bred individuals due to genetic effects and/or plasticity [@weberInteractionsHatcheryWild2003; @snyderLimitationsCaptiveBreeding1996; @arakiHatcheryStockingHelp2010] when considering the reproductive contribution to the next generation. Without loss of generality, we assumed constant $K_i$ ($K_i = K$) and $R_t$ ($R_t = R$) across species and time, respectively.

@@

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the demographic benefit of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across most ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figure S1-6**).

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon (**Figure 1E**), one of Japan's most valuable freshwater resources [@kitadaLessonsJapanMarine2020], on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (**Figure 1C and 1D**), a long-term program exists to monitor stream fish communities along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see **Materials and Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

@@

Competition is a likely mechanism, as our theory assumes. Field experiments confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage, typically one year [@hasegawaWildMasuSalmon2018]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @weberInteractionsHatcheryWild2003]. Importantly, socioeconomic factors control the number of releases (e.g., budget allocation). As such, given the density-dependent regulation of recipient communities, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

@@

We should prioritize habitat preservation and restoration with a broader scope of biodiversity conservation. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of stock enhancement. Protected areas and environmental restoration are promising tools to conserve biodiversity, and the smart spatial design is key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@akasakaSmartProtectedArea2017; @teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Socioeconomic factors such as governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options for the sustainable future exist before blindly accepting stock enhancement. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the stock enhancement will never be effective but impairs biodiversity.

@@

We are grateful to the participants involved in the long-term monitoring program at the protected watersheds in Hokkaido. We thank Masato Yamamichi for helpful discussion on this manuscript. Data and codes are available at <https://github.com/aterui/public-proj_fishery-stability>.

### 2022-02-01 rmd/supporting_info.Rmd
edit: manuscript

**Time-series data.** We assembled time-series fish data across `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. A long-term monitoring program began at these watersheds in 1971, but an effective sampling method (two-pass sampling with a combination of electrofishing and cast net) was implemented in 1999. Most data were collected in summer with irregular interannual intervals (1- to 3-year intervals for most cases), and sampling efforts were quantified by sampling area (average: `r op(mean(df_area$area))` $\pm$ `r op(sd(df_area$area))` m^2^). We confined our analysis to the sites that meet the following criteria: (i) the observation span (from the first to the last year of observation) exceeds 10 years, (ii) the number of observation years exceeds five years, and (iii) masu salmon is observed at least twice during the observation period. As a result, we used time-series data at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds from 1999 to 2019. Summed abundance of first and second passes was used in the following analysis. **Table S5** summarizes observed species in these watersheds.

@@

**Environmental data.** At each sampling site, we measured the following environmental variables as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest), local climates (annual mean air temperature [degree Celsius] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]). We used MERIT Hydro [@yamazakiMERITHydroHighresolution2019] to delineate upstream watershed polygons for each sampling site. We estimated the proportion of forest land use in each watershed polygon based on land-use data in 2015 from Copernicus Global Land Service (100-m resolution) [@marcelbuchhornCopernicusGlobalLand2020]. Climate data at each sampling site were extracted from CHELSA version 1.2 [@kargerClimatologiesHighResolution2017; @kargerDataClimatologiesHigh2018]. Annual data of ocean productivity (2003-2019) were extracted from OceanColor ([\<https://oceancolor.gsfc.nasa.gov/\>](https://oceancolor.gsfc.nasa.gov/){.uri}) and we calculated the average value within the 30-km radius of each river mouth. We used the following R packages to perform geospatial analysis: *sf* [@pebesmaSimpleFeaturesStandardized2018]*, raster* [@hijmansRasterGeographicData2020]*, exactextractr* [@bastonExactextractrFastExtraction2020]*, stars* [@pebesmaStarsSpatiotemporalArrays2020]*, whitebox* [@lindsayWhiteboxGATCase2016]*.*

@@

```{=tex}
\begin{align}
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state}
\end{align}
```
where $ln~r_{s}$ is the site-specific rate of change at site $s$, and $\epsilon_{s,t}^{state}$ is the process error that follows a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the standard deviation $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by partially sharing information across sites [@gelmanDataAnalysisUsing2007]. We used median estimates of fish density $n_{s,t}$ to calculate temporal CVs, means ($\mu$), and SDs ($\sigma$).

@@

We performed a sensitivity analysis of the community simulation to identify key simulation parameters that strongly affect the relationships between community dynamics (temporal mean and SD of density) and stock enhancement. We generated 500 sets of parameter combinations by randomly drawing values of 7 simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we simulated dynamics of 100 independent communities with varying numbers of stock enhancement $R$ (drawn randomly from $Unif(0, 500)$). This yields a total of 50000 simulation replicates. In each simulation replicate, we ran 1600 time steps of community dynamics and obtained temporal means and SDs of the whole community ($\sum_i^SN_i$), enhanced species ($N_1$), and unenhanced species ($\sum_{i, i\ne1}^SN_i$) using the last 1000 time steps. The first 600 time steps were discarded as initialization and burn-in periods.

@@

where $\zeta_k$ ($k = 0-7$) are the intercept ($\zeta_0$) and regression coefficients ($\zeta_{1-7}$). Both response and explanatory variables were standardized to a mean of zero and a standard deviation of one to compare regression coefficients.

@@

  table_legend <- paste0("Sensitivity analysis of the community simulation. Parameter estimates of linear regression models (standard errors in parenthesis) are shown. Response variables are the effects of stock enhancement on temporal mean density ($\\mu$) and SD ($\\sigma$) of ", c_text[i],". Both response and explanatory variables (i.e., simulation parameters) were standardized to a mean of zero and a standard deviation of one before the analysis.")

>>>>

**Time-series data.** We assembled time-series fish data at `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. The Hokkaido Research Organization lead a long-term monitoring program at these watersheds, and the data are published as annual reports [@salmonandfreshwaterresearchinstituteAnnualReportMonitoring2021]. The program began in 1963, but an effective, standardized sampling method was implemented since 1999 (two-pass sampling with a combination of electrofishing and cast net). Most data were collected in summer with irregular interannual intervals (1- to 3-year intervals for most cases), and sampling efforts were quantified by sampling area (average: `r op(mean(df_area$area))` $\pm$ `r op(sd(df_area$area))` m^2^). We confined our analysis to the sites that meet the following criteria: (i) the observation span (from the first to the last year of observation) exceeds 10 years, (ii) the number of observation years exceeds five years, and (iii) masu salmon is observed at least twice during the observation period. As a result, we used time-series data at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds from 1999 to 2019. Summed abundance of first and second passes was used in the following analysis. **Table S5** summarizes observed species in these watersheds.

@@

**Environmental data.** At each sampling site, we measured the following environmental variables as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest), local climates (annual mean air temperature [degree Celsius] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]). We used MERIT Hydro [@yamazakiMERITHydroHighresolution2019] to delineate upstream watershed polygons for each sampling site. We estimated the proportion of forest land use in each watershed polygon based on land-use data in 2015 from Copernicus Global Land Service (100-m resolution) [@marcelbuchhornCopernicusGlobalLand2020]. Climate data at each sampling site were extracted from CHELSA version 1.2 [@kargerClimatologiesHighResolution2017; @kargerDataClimatologiesHigh2018]. Annual data of ocean productivity (2003-2019) were extracted from OceanColor [@seawifsmissionpageNasaOceanBiology2019] and we calculated the average value within the 30-km radius of each river mouth. We used the following R packages to perform geospatial analysis: *sf* [@pebesmaSimpleFeaturesStandardized2018]*, raster* [@hijmansRasterGeographicData2020]*, exactextractr* [@bastonExactextractrFastExtraction2020]*, stars* [@pebesmaStarsSpatiotemporalArrays2020]*, whitebox* [@lindsayWhiteboxGATCase2016]*.*

@@

```{=tex}
\begin{align}
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state}
\end{align}
```
where $ln~r_{s}$ is the site-specific rate of change at site $s$, and $\epsilon_{s,t}^{state}$ is the process error that follows a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the SD $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by partially sharing information across sites [@gelmanDataAnalysisUsing2007]. We used median estimates of fish density $n_{s,t}$ to calculate temporal CVs, means ($\mu$), and SDs ($\sigma$).

@@

We performed a sensitivity analysis of the community simulation to identify key simulation parameters that strongly affect the relationships between community dynamics (temporal mean and SD of density) and stock enhancement. We generated 500 sets of parameter combinations by randomly drawing values of 7 simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we simulated dynamics of 100 independent communities with varying numbers of stock enhancement $R$ (drawn randomly from $Unif(0, 500)$). This yields a total of $5 \times 10^4$ simulation replicates. In each simulation replicate, we ran 1600 time steps of community dynamics and obtained temporal means and SDs of the whole community ($\sum_i^SN_i$), enhanced species ($N_1$), and unenhanced species ($\sum_{i, i\ne1}^SN_i$) using the last 1000 time steps. The first 600 time steps were discarded as initialization and burn-in periods.

@@

where $\zeta_k$ ($k = 0-7$) are the intercept ($\zeta_0$) and regression coefficients ($\zeta_{1-7}$). Both response and explanatory variables were standardized (mean = 0, SD = 1) to compare regression coefficients.

@@

  table_legend <- paste0("Sensitivity analysis of the community simulation. Parameter estimates of linear regression models (standard errors in parenthesis) are shown. Response variables are the effects of stock enhancement on temporal mean density ($\\mu$) and SD ($\\sigma$) of ", c_text[i],". Both response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")
